\chapter{Introduction}
\label{introduction}

This dissertation presents a continuum treatment of growth in
biological tissue developed within the context of modern mixture
theory. The crux of this work is a careful examination of the
assumptions underlying continuum thermodynamics under the condition
that multiple interacting species occupy a region of Euclidean space
simultaneously. The formal axiomatic treatment presented derives from
these assumptions, and provides insight into the sequence of
interaction between tissue mechanics, mass transport and biochemical
reactions. A computational formulation built upon the theory is used
to solve a broad class of numerical examples demonstrating several
biophysical aspects of tissue growth.

This initial chapter provides some context for this work
(Section~\ref{background}) and an overview of the topics considered in
the remainder of the dissertation (Section~\ref{overview}).

\section{Background}
\label{background}

\emph{Growth} involves the addition or depletion of mass in biological
tissue. Growth occurs in combination with \emph{remodelling}, which is
a change in microstructure, and possibly with \emph{morphogenesis},
which is a change in form in the embryonic state. The physics of these
processes are quite distinct, and for modelling purposes can, and
must, be separated. In this work, biological growth is formulated on a
continuum scale within the context of mixture theory
\citep{TruesdellToupin:60, TruesdellNoll:65, BedfordDrumheller:1983},
which allows us to systematically account for the numerous interacting
and inter-converting species constituting the tissue. The crux of this
work is a careful examination of the assumptions underlying continuum
thermodynamics for these mixtures, especially in the presence of
supplementary terms which enhance the balance laws from classical
mechanics to allow for the complex behaviour of tissues.

There have been a number of significant papers on biological growth
(and remodelling), of which we touch upon some whose approaches are
either similar to this work in some respects, or differ in important
ways.

In the context of biological growth, the notion of a mass source was
first introduced in \citet{CowinHegedus:76}. Also recognising the
importance of mass transport to the growth problem,
\citet{EpsteinMaugin:2000} introduced a mass flux term to the
corresponding transport equation. In their work, they also considered
irreversible momentum and entropy contributions from the species flux
to account for these aspects of the inter-species interactions and
deduce non-symmetric partial Cauchy stresses, in contrast to the
treatment here. (See Sections~\ref{balance-of-angular-momentum} and
\ref{eu-balance-of-angular-momentum}.)

\citet{HumphreyRajagopal:02} provided a mathematical treatment of
\emph{adaptation} in a tissue, which includes growth and remodelling
in the sense of this work. They introduced the notion of evolving
natural configurations to model the state of material deposited at
different instants in time. The treatment of the growth part of the
deformation gradient in this work (Section~\ref{kinematics-of-growth})
bears some resemblance to this idea. This concept also forms the basis
for an active active field of study within the literature
\citep{Skalak:81, SkalakHoger:96, Klischetal:2001, TaberHumphrey:2001,
  LubardaHoger:02, AmbrosiMollica:2002} focusing on the kinematic
aspects of on biological growth.

\citet{PreziosiFarina:2002} developed an extension to the classical
Darcy's Law to incorporate mass exchanges between reacting
species. This consideration is relevant to growth problems; however,
these issues were subsumed in \citet{growthpaper}, upon which this
work is based.

While most of the computational examples in this dissertation are
presented in the context of modelling our primary tissues of interest,
engineered tendon and ligament constructs
\citep{Calve:04,Syed-Picard:06}, many of the general ideas presented
in this work are applicable to modelling tumour growth. The ideas
proposed are similar to tumour modelling work that account for
mechanical effects \citep{Jackson-Byrne:02,Byrneetal:06}.

The form of the Clausius-Duhem inequality arrived at in
Section~\ref{eu-entropy-inequality} is equivalent to the forms in
recent work on mixture theory-based models for biological growth
\citep{loret05, ateshian07}. However, subsequently varying choices
made in the different works, including this one, for the constitutive
independent variables result in altered constitutive
specification. Moreover, the constitutive choices detailed in this
work ensure that the Clausius-Duhem inequality is 
satisfied {\em a priori}, are adequately general to handle a fairly
large class of physics, and most significantly, have been implemented
in a coupled formulation retaining much of their rich detail, as
evidenced by the computational examples in
Chapter~\ref{numerical-simulations-2}.

\section{An overview}
\label{overview}

The core of this dissertation is divided into two parts.

The first part, consisting of Chapters~\ref{lagrangian-perspective}
and \ref{numerical-simulations-1}, develops the theoretical
formulation for biological growth from a Lagrangian perspective and
presents representative numerical examples demonstrating aspects of
the coupled physics using a corresponding computational
implementation. This approach, based on our previous work
\citep{growthpaper}, draws in some measure from
\citet{CowinHegedus:76, EpsteinMaugin:2000} and
\citet{TaberHumphrey:2001}, and works in terms of material quantities
defined in the reference configuration of the tissue.

The theoretical treatment presented in Chapters~\ref{lagrangian%
  -perspective} begins by deriving general balance equations governing
the behaviour multi-phase mixtures, and then proceeds to specify
constitutive relationships pertinent to growing biological tissue that
are thermody\-namically-consistent, in the sense that specification of
these relations does not violate the Clausius-Duhem dissipation
inequality. Two important contributions of this work include a
comprehensive account of the coupling between transport and mechanics
(stemming from from the balance equations, kinematics and constitutive
relations), and an improvement to the mathematical treatment that
allows for the numerical stabilisation of the advection-diffusion mass
transport equation in the advection-dominated regime.

This approach was impaired by some basic deficiencies. Firstly, while
the transport equations were posed (consistently) reference
configuration, for a tissue undergoing finite strains, the physics of
fluid-tissue interactions and the imposition of relevant boundary
conditions is best understood and represented in the current
configuration.
Secondly, also stemming from its 
roots in solid mechanics, the formulation relied upon primitive
quantities that are not natural to fluids, such as the {\em
  deformation gradient of the fluid}. While such quantities can be
formally defined, they are not easily tracked during the course of
solving boundary value problems. One final problem with this approach
arose from attempting to impose the balance of momentum for the tissue
as a whole, and this required additional assumptions on the
microstructural mechanics. These require sophisticated homogenisation
techniques (e.g., \citet{IdiartCastaneda:2003}) and have strong implications
for the stiffness of tissue response, the nature of fluid transport,
and since nutrients are dissolved in the fluid, ultimately for
growth. This meant that without additional, complex assumptions, the
formulation could not provide precise, quantitative results. (The calculations in
Section~\ref{constriction-1}, however, do work determine the upper and
lower bounds of the solutions.)

It is these drawbacks that motivate work presented in the second half
of this dissertation, composed of Chapters~\ref{eulerian-perspective} and
\ref{numerical-simulations-2}.

%% Mention the fact that the boundary conditions we work with in the
%% second formulation is much more natural and amenable to
%% experiments.

Chapter~\ref{eulerian-perspective} is a reflection of our current
understanding of the system that has evolved over these past few
years. Recognising that the tissue is composed of many phases that
undergo long-range transport associated with fluids, the formulation
is rederived from an Eulerian perspective; the viewpoint used for
formulating the basic laws of fluid mechanics.

In this approach, the balance equations are derived in the current
configuration of the tissue in terms of spatial variables  enabling a
straightforward application of 
physically-relevant boundary conditions, but more importantly, the
governing equations for the fluid phase are recast in terms of
primitive variables that are more natural to fluid mechanics: fluid
velocity and pressure. Another significant aspect of this work is
that, upon revisiting the Clausius-Duhem inequality in terms of
spatial variables, appropriate constitutive choices are made to ensure
that the inequality is satisfied a priori.

Accompanying this revised formulation is an improved computational
framework, now designed to solve detailed set of balance equations,
eliminating the need for the micromechanics assumptions
mentioned earlier. A discussion of this framework along with selected
computational examples are the subject of
Chapter~\ref{numerical-simulations-2}. This seems promising in that
it demonstrates several basic aspects of the physics of these problems
and has been relatively easy to extend to varied classes of problems,
as evidenced by the examples in Chapter 5. it works well for our
tissues of interest.

The computational framework presents a powerful tool to answer
specific questions—ranging from those pertinent to viscoelastic
aspects of the mechanical response of growing tendons under different
loading conditions, to quantitative investigations of the efficacy of
drugs based on how they are administered, to understanding the
cellular processes associated with tumour growth.

Chapter 6 concludes, As is their very nature, the appendices at the end of this thesis
house topics that while useful, and sometimes interesting, interrupt
the flow of ideas during the main
discourse.

% Local Variables:
% TeX-master: "thesis"
% mode: latex
% mode: flyspell
% End:
